---
title: "R QRG"
output: html_notebook
---
ML & Stats Models
```{r}
#LDA (linear discriminatory analysis
library(MASS)
lda_model <- lda(Group~ acceleration+displacement+horsepower+weight, data=train_4))

#MLR (multiple linear regression)

```

Train and test models
```{r}
#Split dataset into train and test subsets
dt = sort(sample(nrow(autos_4), nrow(autos_4)*.7))
train_4<-autos_4[dt,]
test_4<-autos_4[-dt,]
```

Data importing
```{r}
read.csv(), read.csv2(), read.tsv(), read.delim()
skip = n - skip rows
# A comment I want to skip 
col_names = FALSE
col_names = c("x", "y", "z")
na = "<VALUE OR PLACEHOLDER>"
data.table::fread() - speeeedy table read
parse_*() - character vector -> specialised vector e.g. logical, integer, date
col_*() - corresponding col function for all parse functions
str(parse_date(c("2010-01-01", "1979-10-14")))
problems() - debug parsing function failure
guess_encoding() - UTF-8, Latin1, etc.
guess_parser() - guess character vector
parse_guess() - guess column
problems() and stop_for_problems()

challenge <- read_csv()
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_logical()
  )
)
  challenge2 <- read_csv(readr_example("challenge.csv"), 
    col_types = cols(.default = col_character())
  )
type_convert(df) - convert table with loaded in with character columns
n_max() - 10k-100k (large file), 1k (small file)

read_lines() or read_file() - read into vector of lines or of length 1

Writing a file:
write_*()
write_rds() and read_rds() - from readRDS() and saveRDS() - saved in custom binary format

#Variable types
is.<TYPE>() #check if variable is of type <TYPE>
as.<TYPE>()#convert variable to type <TYPE>

```

Tidy data:
```{r}
Pivoting:
pivot_longer() - rows are columns (increase no. rows)
pivot_wider() - columns are rows (increase no. cols)
e.g. pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
separate() - one col contains multiple variables 
e.g. separate(rate, into = c("cases", "population"), sep = "/")
  convert = TRUE (guess columns types e.g. int, dbl)
    can also use sep = n to split after n characters
unite() - inverse of separate()
select - select columns
separate(key, c("new", "type", "sexage"), sep = "_") - renaming cols
select(-<COL>) - remove column
#Binning continuous variable
a$density_new <- cut(a$density,breaks=c(seq(from = min(a$density),to = max(a$density),length.out=5)))
  #cut breaks into categories
  #Breaks define the break points
  #Seq creates a sequence from defined points
  #Length.out defines length of sequence
```

Relational data and joins:
```{r}
#Setting names for a data frame
a <- data.frame(row.names = "mpg" %+% chr(1:nrow(mpg)),
  manufacturer = mpg$manufacturer,
  model = mpg$model, displ = mpg$displ,
  cyl = mpg$cyl)

#Adding observations
total <- rbind(dataframeA, dataframeB)
      #ensure number of variables match (can add col with NAs or delete extra cols not intersecting)
#Example
grp <- runif(nrow(custdata))
  df <- data.frame(grp)
  newdata <- cbind(custdata, df)
  newdata["grp"]

#Joins - ensure variable names are identical to join
Mutating joins - add new vars of one table to another
Filtering joins - filter observations from one table based on another
Set operations - treat observations as if they were set elements
Inner join - keep observations that appear in both tables
Outer join - keep observations that appear in AT LEAST 1 table

#Mutating joins - add new variables to one data frame from matching observations in another
     left_join(<TABLE1>, <TABLE2>, by=<KEY>) - keep all in 1
     right_join(<TABLE1>, <TABLE2>, by=<KEY>) - keep all in 2
     inner_join(...) - keep intersecting cols 
     full_join(...) - keep all cols
     by = NULL - variables appearing in both tables (natural join)
     by = c("a" = "b") - key has different name in tables
     
#Filtering joins - filter observations from one data frame based on whether or not they match an observation in the other table
      semi_join(x,y) - keep all in x that have a match in y
      anti_join(...) - drops all in x that have a match in y
      flights %>% 
        filter(dest %in% top_dest$dest) - filter flights by rows in dest col from top_dest table 
          OR flights %>% 
              semi_join(top_dest)
          
#Set operations - treat observations as if they were set elements
      intersect(x, y): return only observations in both x and y.
      union(x, y): return unique observations in x and y.
      setdiff(x, y): return observations in x, but not in y.
          
     
#Keys:
key - variable connecting tables
primary key - uniquely identifies an observation in its own table
foreign key - uniquely identifies an observation in another table
surrogate key - artificial key using mutate() & row_number()
Checking primary keys - count() -> n=1
```

Strings:
```{r}
**See 14.4.1 for string-based col filtering**
**Replace 'str' with 'stri' for stringi instead of stringr**
Contains string:
  df$loc01 <- ifelse(grepl("non",df$loc_01),'outside','inside')
str_replace(<COL>,"<STRING>","<REPLACMENT>")
str_length() and other str_*()
str_c() - combine strings 
  sep = '<SEPARATOR>' 
  str_replace_na(x) - replace missing values with NA and combine
  str_c("prefix-", c("a", "b", "c"), "-suffix") - prefix and/or suffix all values in middle vector and store as a vector
**Can use logic operators to include prefix/suffix** (14.2.2)
**Specify locale (region) if needed** (14.2.4)
str_sub(<STRING>, start, end) 
str_to_lower() - convert to lower case
       upper or title - upper case or title
str_view() / str_view_all() - take char vector and reg expression and show how they match
  x <- c("apple", "banana", "pear")
    str_view(x, "an")
    or str_view(x, ".a.") - match any character
      "\\." when trying to match a "." (dot)
      "\\\\" when trying to match a "\"
      "^<TO MATCH> - match start of string
      <TO MATCH>$ - match end of string
      ^<TO MATCH>$ - match complete string
      \d: matches any digit.
      \s: matches any whitespace (e.g. space, tab, newline).
      [abc]: matches a, b, or c.
      [^abc]: matches anything except a, b, or c.
      ']' '\' '^' and '- have to be handled with double backslash
      escapes
      str_view(c("grey", "gray"), "gr(e|a)y")
    Specify number of matches e.g. str_view(x, "C{2}")
      {n}: exactly n
      {n,}: n or more
      {,m}: at most m
      {n,m}: between n and m
      {n,m}: between n and m - matching shortest string
str_detect() - determine if character vector matches a pattern
  uses logical operators --> incorporate sum and mean (0,1)
str_count() - outputs no. of matches
  **Useful with mutate (14.4.1)**
str_subset() - wrapper???
str_view_all() - show all matches, not just the first one
str_extract() - extract actual text of a match
str_extract_all() - extract all actual text matches
  simplify = TRUE to return matrix with short matches expanded to the same length as the longest (consistency)
str_match() - gives each individual component, returning a matrix (one col for complete match followed by one col for each group)
str_replace() - replace matches with new strings (first)
str_replace_all() - replace ALL matches with new strings
  e.g x <- c("1 house", "2 cars", "3 people")
        str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
str_split() - split into pieces using delimiter
  simplify = TRUE to transform into matrix
    e.g. fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
            %>% str_split(": ", n = 2, simplify = TRUE)
    boundary() - split by char, line, sentence or word
str_locate() - starting & ending position of each match
      or  _all()
writelines() - raw contents of string
ignose_case = TRUE - match either upper or lower case forms 
multiline = TRUE - allows ^ and $ to match start and end of each line, instead of complete string e.g. when using \n (new line)
comments = TRUE
dotall = TRUE - allow "." (dot) to match everything (inc \n)
regex() - used by default 
  Alternatives:
      fixed() - match exactly specified sequence of bytes
      coll() - compare strings using standard collation rules 
          Note: coll() is slower
apropos() - searches all objects available in global environment
  e.g. apropos("replace") --> "%+replace%"       "replace"          "replace_na"
dir(pattern = "<PATTERN>") - list all files in a directory
"\n" - new line()
"\t" - tab
Non-English - "\u00b5" --> Greek mew
```

Factors:
```{r}
**Factors are used to work with categorical variables, variables that have a fixed and known set of possible values. They are also useful when you want to display character vectors in a non-alphabetical **
**Modifying factor order - 15.4**
  
Create list of valid 'levels'
e.g. month_levels <- c(
      "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
      "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
factor() - normal factoring
parse_factor() - give warnings if not in levels list
  y1 <- factor(<TABLE>, levels = month_levels) 
sort() - sort based on levels sorting
unique() - order of levels match first appearance, not alphabetical
fct_inorder() - same as unique, but separate, after-the-fact fnc
fct_reorder() - input factor (f), numeric value used for ordering (x)
  fun - if there are multiple values of x 
fct_relevel(<col>, <level1>,<leveln> ) - move levels to the front
fct_reorder2() - reorders factors by the y values associated with the largest x values (line plots)
fct_infreq() - bar plots - order levels in increasing freq
  fct_rev() - combine with above (for some reason)
fct_recode() - recode/change value of each level
  e.g gss_cat %>%
        mutate(partyid = fct_recode(partyid,
          "Republican, strong"    = "Strong republican",
          "Republican, weak"      = "Not str republican",
          "Other"                 = "No answer",
          "Other"                 = "Don't know",
          "Other"                 = "Other party" ... ))
fct_collapse() - same as recode but assign using vectors
fct_lump() - specify n groupings and lump into categories
```


Arithmetic, vectors, matrices, stats
```{r}
c() - vector
n() - returns size of current group
%/% - integer division (5.5)
%% - remainder

Cumulative and rolling aggregates - cumsum(), cumprod(), cummin(), cummax(), cummean()
Compute running differences (5.5) - lead(x), lag(x)

%>% - Pull outcome of previous action into next lines function to avoid having to define new variables constantly e.g. x %>% f(y) turns into f(x, y) 

count(), mean(), sum(), median(), min(), max(), quantile(), sd(), IQR(), mad(), first(), nth(x,n), last()

Aggregation, NAs, weight
n() - return size of current group
sum(!is.na(x)) - count non-NA values
n_distinct(x) - count distinct values
count(x,wt = <VAR>) - count <VAR> for grouping x
sum(x > 500) - TRUE values = 1 -> sum all 1s
5.6.5 for progressively rolling up summaries

<FUNCTION>(par()) #determine number of parameters in a function
par(<PARAMETER>) #default of parameter

#Subset rows based on variable condition
a[a$manufacturer == "audi" & a$model == "a4 quattro",]
```

Date and time
```{r}
as.Date(x, "input_format")

#Symbol Meaning Example
%d Day as a number (1-31) 01-31
%a Abbreviated weekday Mon
%A Unabbreviated weekday Monday
%m Month (1-12) 01-12
%b Abbreviated month Jan
%B Unabbreviated month January
%y 2-digit year 18
%Y 4-digit year 2018
Default - yyyy-mm-dd

#Formatting date values
today <- Sys.Date()
format(today, format="%B %d %Y")

#Date difference
julian(<DATE1>, origin = <DATE2>)
difftime(<DATE1>, <DATE2>, units="weeks")

Sys.Date() - current system date in date type
date() - current system data in character/string type
Sys.time() contains timezone, of class c("POSIXct" "POSIXt")
weekdays()
quarters()
today() - date
now() - date and time and timezone
ymd, mdy, dmy ("<STRING>") --> parse
make_date() or make_datetime() - create date/time from multiple cols that are individually month, year, time, etc.
as_datetime() or as_date() - convert between date and date/time
year(), month(), mday(), yday(), wday(), hour(), minute(), second()
  label = TRUE = return abberviated name of the month or day of the week 
  abbr = FALSE - return full name

Rounding to nearest unit of time:
floor_date() - round down
round_date() - choose
ceiling_date() - round up

Setting values:
year(datetime) <- 2020
month(datetime) <- 01
hour(datetime) <- hour(datetime) + 1
update(datetime, year = 2020, month = 2, mday = 2, hour = 2)

Durations (16.4.1) - exact no. of seconds
**Can add, multiply, subtract results**
h_age <- today() - ymd(19791014)
 > Time difference of 14817 days
as.duration(h_age)
 > [1] "1280188800s (~40.57 years)"
dseconds(), dminutes() ... dyears()

Periods (16.4.2) - human time periods e.g. weeks/months
**Can add, multiply, subtract results**
seconds, minutes ... years()
ymd("2016-01-01") + years(1)
 > [1] "2017-01-01"

Intervals (16.4.3) - duration with start/end points
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)
 > [1] 365


```


Logic operators and if statements
```{r}
< =< != == &(and) |(or) %in% (subset)
#if and else statement
if (logical_expression) {
statements
} else {
alternative_statements
#ifelse
ifelse (logical_expression, yes_statement, no_statement)
}

#Create new column based on logical expressions and existing columns 
autos_4 <- mutate(autos_4, Group = case_when(mpg < 20 ~ "low", mpg < 27 ~ "medium",mpg >= 27 ~ "high"))

```

Looping (for and while)
```{r}
#For
for(i in x:y) {
print(i*i)
}

#While
i <- 1
while(i<=5) {
print(i*i)
i<-i+sqrt(i)
}

```

Summaries
```{r}
view(), str(), summarise()
dim() - table dimensions
group_by(dataframe, col1, col2)
e.g. summarise(group_by(col1, col2),col3)
--> aggregate col3 by col1 and col2
ungroup() - remove aggregation
range(<col>) #numeric range of col
```


dplyr key functions (data frames and tables)
```{r}
filter()
arrange()
select()
mutate(), transmute()
summarise()
group_by()
rename() --> variable = var_new
  #custdata <- rename(custdata, age.cat=agecat)

```

Mutate
```{r}
mutate(DATAFRAME,new_col = col1 - col2)
Note: use transmute to keep old columns (col1, col2)
```

Functions
```{r}
#General function layout
add <- function(<input1>,<input2>) {
result <- input1+input2
return(result)
}
```

Filtering and selecting
```{r}
#Filtering - selecting rows
filter(DATAFRAME,VARIABLE==1)
OR
newdata <- custdata[which(custdata$sex=="M" & custdata$age < 30),] #col indices blank to include all cols, just dropping rows (filtering)

#Selecting or subsetting - selecting cols
select(DATAFRAME,VARIABLE_1,VARIABLE_5) --> only shows columns specified
Note: use '-' to exclude specified rows
OR
myvars <- c("custid", "is.employed", "income",
"marital.stat", "health.ins", "age") #specify cols to grab later in vector form
newdata <- custdata[myvars] #call cols to keep from character vector of col names
OR #not include
myvars <- names(custdata) %in% c("sex", "state.of.res")
newdata <- custdata[!myvars]
OR #explicit list of cols
newdata <- subset(custdata, age >= 65 | age < 24,
select=c("custid", "marital.stat"))
OR #from colx to colx+n
newdata <- subset(custdata, sex=="M" & age < 25,select=custid:age)

#Helpful functions
starts_with(), ends_with(), contains(), matches(), everything()
```

Arranging, ordering, ranking, sorting
```{r}
arrange(DATAFRAME,VARIABLE_1,VARIABLE_2)
desc() - descending
Note: missing values at end

min_rank(x) (desc(x)) for reverse
Variants: row_number(), dense_rank(), percent_rank(), cume_dist()

#Sorting data
newdata <- custdata[order(<COL1>, -<COL2>),] #negative for descending order
                                              #sorting by asc <COL1> and then desc by <COL2> 

```

Missing/NA values
```{r}
 df[!(is.na(df$start_pc) | df$start_pc==""), ] #Remove records with nulls
is.na(x)
na.rm = TRUE - remove NAs before computing function
Replace with NA - mutate(y = ifelse(y < 3 | y > 20, NA, y))
**An explicit missing value is the presence of an absence; an implicit missing value is the absence of a presence.**
  -> can use pivot_*(), complete() and fill() to make implicit explicit
  -> values_drop_na = TRUE in pivot_longer() to make explicit implicit
str_replace_na()
```

Sample data or test data
```{r}
#Creating sample data
mysample <- custdata[
sample(1:nrow(custdata), 3, replace=FALSE),]
```


Plotting
```{r}
#ggplot
Required:
ggplot(data=<DATA>)+
      <GEOM_FUNCTION>(mapping=aes(x=col1,y=col2),

Additional - part of mappings
      stat=<STAT>,position=<POSITION>) +

Additional - functions
      <COORDINATE_FUNCTION> +
      <FACET_FUNCTION> +
      <SCALE_FUNCTION> +
      <THEME_FUNCTION>

  Examples:
    coord_cartesian(ylim = c(0,50))
    Note: xlim(), ylim() diff b/w coord_cartesian 
    and ggplot2 functions 

Labels:
  labs(title = "Fuel efficiency generally decreases with engine size")
    title, subtitle, caption, x, y, colour
Mathematical labels: x = quote(<EQUATION>)
Point labelling:
     geom_text(aes(label = model), data = best_in_class)
  or geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
  or ggrepel::geom_label_repel(aes(label = model), data = best_in_class)
geom_hline() or _vline() - reference lines
geom_rect() - draw rectangle around points of interest
geom_segment() - draw attention to a point with an arrow
geom_smooth(method="lm" or "loess") #methods of line/curve fitting - regression or local regression
  #Level = 0.95 (confidence)
    
Legend:
theme(legend.position = "none" - turn off legend
base + theme(legend.position = "left") 
  **left, top, bottom, right (default), none**
      
Scales (see 28.4.3 for palettes):
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_colour_discrete()
      **continuous, discrete, datetime, date**
Axis ticks: scale_y_continuous(breaks = seq(15, 40, by = 5))
guides(), guide_legend(), guide_colourbar()
Changing scale (note 'log')
    ggplot(diamonds, aes(carat, price)) +
      geom_bin2d() + 
      scale_x_log10() + 
      scale_y_log10()
Assign colours to specific values: scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
Continuous colour - scale_colour_gradient() or scale_fill_gradient()
  or viridis::scale_colour_viridis()
Diverging continuous scale - scale_colour_gradient2()
Categorical - scale_colour_brewer(palette = "Set1")

Zooming:
coord_cartesian(xlim = c(5, 7), ylim = c(10, 30))

Themes:
theme_*()

#Geoms

#Aesthetics

Aesthetics (see 28.4.3 for palettes):
size = <col1>
se = TRUE/FALSE (geom_smooth) #se = standard errors (confidence intervals)
linetype = ...
colour, shape, fill, size
+facet_wrap(facets=a$quality)
+facet_grip(.~<COL>) # "." indicates single row and "col" columns (number of unique values in col)
facets=.~<col> # "." indicates single row and "col" columns (number of unique values in col)
binwidth=n
+ggtitle("MAIN TITLE")
+bw_theme(base_family="FONT")
+ylim(-3,3) #limit data
+coord_cartesian(ylim=c(-3,3)) #Limit coordinate system
+labs(x,y,title)
+annotate("text", x = 150000, y = 3000, label = "<LABEL>")
#Log
+scale_x_log10(breaks=c(100,1000,10000,100000) # or y - change scale not values

Simplified:
  ggplot(faithful, aes(eruptions)) + 
    geom_freqpoly(binwidth = 0.25)

#Categorical:
  Single variable:
  ggplot(data=<DATA>) +
    geom_bar(mapping=aes(x=<VAR>),position="")
    #position = dodge, stack, fill, identity
Manual: count(<VAR>)

#Continuous:
  Single variable:
  ggplot(data=<DATA) +
    geom_histogram(mapping=aes(x=<VAR>,binwidth=n))
or  geom_freqpoly(...) - line
Manual: count(cut_width(<VAR>,n))
  
  #Single variable broken down by categorical 
  A) Standardise - density - area under curve equals 1
    ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)
    OR
  B) Boxplots
    ggplot(data = diamonds, mapping = aes(x = cut, y = price)) +
  geom_boxplot() 
        #Boxplot parameters - outlier.colour="red", outlier.shape=16, outlier.size=2
    
#Two categorical variables:
  A) Geom count (dot sizes)
  ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))
  B) Geom tile (hot spots)
  diamonds %>% 
    count(color, cut) %>%  
    ggplot(mapping = aes(x = color, y = cut)) +
      geom_tile(mapping = aes(fill = n))

#Two continuous variables:
  A) Scatter plot (use alpha to add heat effect)
  ggplot(data = diamonds) +
    geom_point(mapping = aes(x = carat, y = price, alpha = 1/100))
  B) 2d/hex binning
  ggplot(data = smaller) +
    geom_hex(mapping = aes(x = carat, y = price))
 or geom_bin2d(...)

Useful functions within ggplot:
  reorder()
  coord_flip()
    ,reorder = (x=COUNT_ACCIDENT))
  
  
Save plots:
  ggsave

Figure sizing:
  fig.width(), fig.height(), fig.asp(), out.width(), out.height()
  fig.width = 6
  fig.asp = 0.618
  out.width = "70%" 
  fig.align = "center"
Multiple
  out.width = "50%" (2), "33%" (3), "25%" (4) ... etc.
  fig.align = "default"

fig.show = "hold" - plots are shown after code
fig.cap 

#Pairwise variable comparison
library(lattice)
pairs(<DATA>)

event_palette <- RColorBrewer::brewer.pal(n=5,'Set2')

#Density plots
hist(data$column,freq=FALSE)

#Boxplot all variables with frequencies
ggplot(stack(<DATA>)) +
+         geom_boxplot(mapping=aes(x=ind,y=values))

#Lattice package
xyplot (scatter), bwplot (box & whisker), and histogram
y~x (y depends on x)
xyplot(y ~ x | f * g, <data>) #f & g are optional conditioning variables

#R default plotting
boxplot(<COLUMN>)
with(<DATA>,plot(X,Y))
abline(h=12,v=2) - draw vert or hori lines on chart
hist(<COL>, breaks = n)  breaks = bins
rug(<COL>) - representation of actual data points under histogram (good for distribution)
barplot()
boxplot(y~x,<DATA>) - boxplot distribution of y column values aggregated by x column as x-axis
with(<DATA>,plot(x,y))
title(main="Old Faithful Geyser data")
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4, 1))
  # | for facets, layout specifying plots cols and rows
base: plot,hist,boxplot,barplot, etc
grDevices: x11, PDF, PostScript, PNG, etc
  xlab, ylab, col.axis, col.lab
  legend("topright",pch=c(17,8),col=c("blue","red"),legend=c("May","Other Months")) #legends, pch=shape (for two subsets, may (blue) and notmay (red))
  lty = line type, lwd = width
  
#COLOURS!!!
#grDevices package
colors() #657 predefined colors
colorRamp() #for palettes
colorRampPalette() #for palettes

#RColorBrewer
cols <- brewer.pal(n,"BuGn") #n colours, palette code
pal<-colorRampPalette(cols)


```

Modelling
```{r}
Building a model:
~~~
  library(modelr)

  mod <- lm(log(price) ~ log(carat), data = diamonds)

  diamonds2 <- diamonds %>% 
    add_residuals(mod) %>% 
    mutate(resid = exp(resid))

  ggplot(data = diamonds2) + 
    geom_point(mapping = aes(x = carat, y = resid))
~~~
```

R packages
```{r}
tidyverse - set of packages working together
d3heatmap or heatmaply - interactive heat plots
hexbin - rectangular binning
modelr - models and predictions
feather - fast binary file format 
ggplot2 - advanced plotting
dplyr - grammar of data manipulation (similar to SQL)
tidyr - tidy data
stringr - tidyverse string manipulation
stringi - tidyverse string manipulation (advanced)
forcats - tools for categorical variables (factors)
ggrepel - auto adjust labels to not overlap
RColorBrewer - palettes (categorical)
viridis - continuous analog of categorical ColorBrewer
ggthemes - themes
lattice - pairwise variable comparison
sqldf - SQL
```

Extras and drafts
```{r}
accident_by_classification <- injury %>% 
        group_by(CLASSIFICATION) %>% 
        summarise(ACCIDENT_TYPE = count(ACCIDENT_DT),na.rm = T)
accident_by_classification <- transform(accident_by_classification, CLASSIFICATION=reorder(CLASSIFICATION, ACCIDENT_TYPE))

ggplot(accident_by_classification, mapping=aes(x=CLASSIFICATION, y=ACCIDENT_TYPE))+
  geom_bar() +
  coord_flip()
  theme_light()


aggregate(x= injury$DOCUMENT_NO, FUN = count, by = list(injury$OPERATOR_NAME))

#Adding zeroes to 2400 time
   ifelse(grepl("INJURIES",injury$ACCIDENT_TIME) | grepl("DAYS",injury$ACCIDENT_TIME) | grepl("DYS",injury$ACCIDENT_TIME),'Injury',
                  ifelse(grepl("ACCIDENT",injury$ACCIDENT_TIME),'Accident only',
                  ifelse(grepl("PERM",injury$ACCIDENT_TIME),'Permanent disability',
                         'Other')))))

ggplot(exper_events,aes(x=EXPER_BIN,y=DAYS_LOST))+
  geom
```
